---
title: "Bcell_extraction_Azimuth"
author: "Daiki Takeuchi, Yoshiaki Yasumizu"
date: "2022/04/23"
output: html_notebook
---

# Ensure

```{r Ensure, eval=FALSE, include=FALSE}
# Ensure Seurat v4.0 or higher is installed
if (packageVersion(pkg = "Seurat") < package_version(x = "4.0.0")) {
  stop("Mapping datasets requires Seurat v4 or higher.", call. = FALSE)
}

# Ensure glmGamPoi is installed
if (!requireNamespace("glmGamPoi", quietly = TRUE)) {
  if (!requireNamespace("BiocManager", quietly = TRUE)) {
    BiocManager::install("glmGamPoi")
  }
}

# Ensure Azimuth is installed
if (packageVersion(pkg = "Azimuth") < package_version(x = "0.3.1")) {
  stop("Please install azimuth - remotes::install_github('satijalab/azimuth')", call. = FALSE)
}
```

# 1. Library

```{r library}
library(Seurat)
library(SeuratData)
library(Azimuth)
library(patchwork)
library(tidyverse)
library(sctransform)
```

# 2. Parameter setting

```{r parameter setting}
project.name <- c("GSE157278")
select.B.meta.data <- c("orig.ident", "Cell", "sample", "disease", "disease_duration",
                        "condition", "age", "sex", "immunosuppressant", "immunosuppressant_duration",
                        "batch")
genes <- c("CD79A", "MS4A1")
prefix_data <- paste0("../../data/", project.name, "/")
prefix_output <- paste0("../../output/", project.name, "/", project.name)
```

# 3. Load the reference

```{r Load_the_reference}
reference <- LoadReference(path = "https://seurat.nygenome.org/azimuth/references/v1.0.0/human_pbmc")
```

# 4. Reading data(query object)

## 4-1. Read10x

```{r reading data}
query.data <- Read10X(data.dir = prefix_data)
query <- CreateSeuratObject(counts = query.data,
                            project = project.name,
                            min.cells = 3,
                            min.features = 200)
query
```

```{r pbmc@assays}
query@assays[["RNA"]]@counts[1:10,1:20]
```

## 4-2. Loading batch_file and metadata

```{r loading batch_file and metadata}
batch <- read_table(paste0(prefix_data, "/cell_batch.tsv.gz"))

batch_select <- query@assays[["RNA"]]@data@Dimnames[[2]] %>%
  tibble()
colnames(batch_select) <- c("Cell")
batch <- right_join(batch,
                    batch_select,
                    by = "Cell")

batch <- batch %>%
  mutate(sample = batch) %>%
  mutate(disease = recode(sample,
                          "HC-1" = "healthy",
                          "HC-2" = "healthy",
                          "HC-3" = "healthy",
                          "HC-4" = "healthy",
                          "HC-5" = "healthy",
                          "pSS-1" = "sjögren's syndrome",
                          "pSS-2" = "sjögren's syndrome",
                          "pSS-3" = "sjögren's syndrome",
                          "pSS-4" = "sjögren's syndrome",
                          "pSS-5" = "sjögren's syndrome")) %>%
  mutate(disease_duration = NA) %>%
  mutate(condition = recode(sample,
                            "HC-1" = "healthy",
                            "HC-2" = "healthy",
                            "HC-3" = "healthy",
                            "HC-4" = "healthy",
                            "HC-5" = "healthy",
                            "pSS-1" = "primary sjögren's syndrome",
                            "pSS-2" = "primary sjögren's syndrome",
                            "pSS-3" = "primary sjögren's syndrome",
                            "pSS-4" = "primary sjögren's syndrome",
                            "pSS-5" = "primary sjögren's syndrome")) %>%
  mutate(sex = recode(sample,
                      "HC-1" = "female",
                      "HC-2" = "female",
                      "HC-3" = "female",
                      "HC-4" = "female",
                      "HC-5" = "female",
                      "pSS-1" = "female",
                      "pSS-2" = "female",
                      "pSS-3" = "female",
                      "pSS-4" = "female",
                      "pSS-5" = "female")) %>%
  mutate(age = NA) %>%
  mutate(immunosuppressant = NA) %>%
  mutate(immunosuppressant_duration = NA) 
  
batch <- batch %>%
  relocate(Cell,
           sample,
           disease,
           disease_duration,
           condition,
           age,
           sex,
           immunosuppressant,
           immunosuppressant_duration,
           batch)

rownames(batch) <- batch$Cell

query <- AddMetaData(query,
                     metadata = batch)
```

# 5. Preprocessing

```{r preprocessing}
query <- PercentageFeatureSet(query,
                              pattern = "^MT-",
                              col.name = "percent.mt")
head(query@meta.data, 5)
```

```{r QC_ViolinPlot}
options(repr.plot.width = 8, repr.plot.height = 5)

VlnPlot(query,
        features = c("nFeature_RNA", "nCount_RNA", "percent.mt"),
        ncol = 3,
        pt.size = 0)
ggsave(paste0(prefix_output, "_check_QC_vlnPlot_B_extraction.pdf"),
       width = 5,
       height = 10)
```

```{r QC_subset}
query <- subset(query,
                subset = nFeature_RNA > 200 & percent.mt < 10)
```

# 6. SCTransform Normalization

```{r sctransform_normalization}
query <- SCTransform(object = query,
                     assay = "RNA",
                     new.assay.name = "refAssay",
                     residual.features = rownames(x = reference$map),
                     reference.SCT.model = reference$map[["refAssay"]]@SCTModel.list$refmodel,
                     method = "glmGamPoi",
                     ncells = 2000,
                     n_genes = 2000,
                     do.correct.umi = FALSE,
                     do.scale = FALSE,
                     do.center = TRUE)
```

# 7. Find anchors between query and reference

```{r Find anchors between query and reference}
anchors <- FindTransferAnchors(reference = reference$map,
                               query = query,
                               k.filter = NA,
                               reference.neighbors = "refdr.annoy.neighbors",
                               reference.assay = "refAssay",
                               query.assay = "refAssay",
                               reference.reduction = "refDR",
                               normalization.method = "SCT",
                               features = intersect(rownames(x = reference$map),
                                                    VariableFeatures(object = query)),
                               dims = 1:50,
                               n.trees = 20,
                               mapping.score.k = 100)
```

# 8. Transfer cell type labels and impute protein expression

```{r Transfer cell type labels and impute protein expression}
refdata <- lapply(X = "celltype.l1",
                  function(x) {
                    reference$map[[x, drop = TRUE]]
                    })
names(x = refdata) <- "celltype.l1"
if (TRUE) {
  refdata[["impADT"]] <- GetAssayData(
    object = reference$map[['ADT']],
    slot = 'data'
  )
}

query <- TransferData(reference = reference$map,
                      query = query,
                      dims = 1:50,
                      anchorset = anchors,
                      refdata = refdata,
                      n.trees = 20,
                      store.weights = TRUE)
```

# 9. Calculate the embeddings of the query data on the reference SPCA

```{r Calculate the embeddings of the query data on the reference SPCA}
query <- IntegrateEmbeddings(anchorset = anchors,
                             reference = reference$map,
                             query = query,
                             reductions = "pcaproject",
                             reuse.weights.matrix = TRUE)
```

# 10. Calculate the query neighbors in the reference with respect to the integrated embeddings

```{r Calculate the query neighbors}
query[["query_ref.nn"]] <- FindNeighbors(object = Embeddings(reference$map[["refDR"]]),
                                         query = Embeddings(query[["integrated_dr"]]),
                                         return.neighbor = TRUE,
                                         l2.norm = TRUE)
```

# 11. NNTransform correcting the Neighbors to account for the downsampling

```{r NNTransform}
query <- Azimuth:::NNTransform(object = query,
                               meta.data = reference$map[[]])
```

# 12. Project the query to the reference UMAP
```{r Project the query to the reference UMAP}
query[["proj.umap"]] <- RunUMAP(object = query[["query_ref.nn"]],
                                reduction.model = reference$map[["refUMAP"]],
                                reduction.key = 'UMAP_')

```

# 13. Calculate mapping score and add to metadata

```{r Calculate mapping score and add to metadata}
query <- AddMetaData(object = query,
                     metadata = MappingScore(anchors = anchors),
                     col.name = "mapping.score")
```

# 14. Visualizations

```{r predicted metadata field}
id <- "celltype.l1"
predicted.id <- paste0("predicted.", id)

# DimPlot of the reference
#DimPlot(object = reference$plot,
#        reduction = "refUMAP",
#        group.by = id,
#        label = TRUE) + NoLegend()

# DimPlot of the query, colored by predicted cell type
DimPlot(object = query,
        reduction = "proj.umap",
        group.by = predicted.id,
        label = TRUE) + NoLegend()
ggsave(paste0(prefix_output, "_check_dimplot_B_extraction.pdf"),
       width = 5,
       height = 10)

## Plot the score for the predicted cell type of the query
#FeaturePlot(object = query,
#            features = paste0(predicted.id, ".score"),
#            reduction = "proj.umap")
#VlnPlot(object = query,
#        features = paste0(predicted.id, ".score"),
#        group.by = predicted.id,
#        pt.size = 0) + NoLegend()

# Plot the mapping score
#FeaturePlot(object = query,
#            features = "mapping.score",
#            reduction = "proj.umap")
#VlnPlot(object = query,
#        features = "mapping.score",
#        group.by = predicted.id,
#        pt.size = 0) + NoLegend()
```

# 15. Check

## 15-1. Dot Plot

```{r DotPlot for finding CD4+Tcell}
options(repr.plot.width = 10, repr.plot.height = 8)

DotPlot(object = query,
        features = genes,
        group.by = predicted.id) + RotatedAxis()
ggsave(paste0(prefix_output, "_check_dotplot_B_extraction.pdf"),
       width = 5,
       height = 10)
```

## 15-2. Feature Plot

```{r FeaturePlot for finding CD4+Tcell}
options(repr.plot.width = 10, repr.plot.height = 8)

FeaturePlot(object = query,
            features = genes,
            reduction = "proj.umap",
            ncol = 2,
            raster = TRUE)

ggsave(paste0(prefix_output, "_check_featureplot_B_extraction.pdf"),
       width = 5,
       height = 10)
```

# 16. Extraction of Bcell

```{r Extraction of CD4+Tcell}
pbmc_B <- subset(query,
                 predicted.celltype.l1 == "B")

pbmc_B.meta.data <- pbmc_B@meta.data
pbmc_B.meta.data <- pbmc_B.meta.data %>%
  select(all_of(select.B.meta.data))

pbmc_B <- GetAssayData(object = pbmc_B,
                       slot = "data")
```

# 17. saveRDS

```{r saveRDS}
saveRDS(pbmc_B.meta.data,
        file = paste0(prefix_output, "_B_MetaData.rds"))

saveRDS(pbmc_B,
        file = paste0(prefix_output, "_B_AssayData.rds"))
```

# 18. Save workspace

```{r save.image}
save.image(file = paste0(prefix_output, "_220423_Bcell_extraction_Azimuth.RData"))
```

# 19. sessionInfo

```{r sessionInfo}
sessionInfo()
```
